### 2016-02-11 inputs/results.tex
several adds after afternoon pass

\subsection{Genotypes}
\label{sec:genotypes}  
GBS detected 32\,706, 40\,734 and 166\,418 SNP loci in Alfalfa-PV, Alfalfa-Med and Rice, respectively (Table~\ref{tab:descriptive_statistics}). The overall missing rate was $66.6\%$, $59.6\%$ and $53.4\%$ in the three datasets. Supplementary Figure S1 shows the density distribution of missing rate per markers in the complete datasets. The amount of missing marker genotypes varied with the threshold of maximum per-marker missing-rate allowed in the data (10\%, 20\%, 40\% and 70\%) and the proportion of artificial missing genotypes that were introduced (1\%, 5\%, 10\% and 20\%). Missing data point counts ranged from a minimum of 3\,878 in Alfalfa-PV with 10\% allowed and 1\% artificial missing genotypes, to a maximum of 5\,580\,616 in Rice with 70\% allowed and 20\% artificial missing genotypes. Over all datasets and allowed/artificial missing genotypes thresholds, the amount of missing data points to be imputed averaged 249\,405.\\
Minor Allele Frequency (MAF) was 0.172, 0.170 and 0.140 in Alfalfa-PV, Alfalfa-Med and Rice, respectively. The three datasets were therefore unbalanced with respect to genotype classes: Rice had the lowest MAF, while Alfalfa-PV and Alfalfa-Med had the smallest minor classes (minor homozygotes 3.5\% and 3.2\% of the total genotypes). Figure~\ref{fig:genotype_classes} reports the proportion of genotypes in each class in the complete datasets (100\%) and as a function of the maximum allowed missing rate per marker. The relative proportion of genotype classes remained approximately stable in rice. In contrast, the heterozygous class in alfalfa tended to become smaller with increasing proportion of allowed genotypes. This reflected the different GBS calling criteria for homozygous and heterozygous SNP: heterozygous SNPs required fewer overlapping reads to be called, while the implemented quality filters on alfalfa required a larger number of reads to call a homozygous loci. Therefore homozygous loci tended to have higher missing rates, and to
be included progressively more with more relaxed thresholds on allowed missing level.
The average missing-rate and MAF in the 12 rice chromosomes were comparable to the whole-rice dataset.\\
Alignment on reference genome assigned a position to 88.66\%, 57.54\% and 57.86\% of Rice, Alfalfa-Med and Alfalfa-PV markers, respectively.

@@

In rice, the average imputation accuracy over all genotypic classes was above 80\% for all methods. No significant difference was found between chromosomes. Thus average accuracy over all chromosomes is therefore presented (Figure~\ref{fig:accuracy_rice}). \\
The overall accuracy for alfalfa, averaged across the two datasets, was 12-25\% lower than rice, ranging between 66\% and 84\%. No significant difference were found between the two datasets. Thus average accuracy over the Alfalfa-Med and Alfalfa-PV is presented (Figure~\ref{fig:accuracy_alfalfa}).

The imputation accuracy varied across the different genotype classes, with the highest accuracy in the most common genotype class (averaging 92.27\% and 96.47\% in alfalfa and rice, respectively), while the least common genotype class showed much lower accuracy (averaging 5.79\% and 69.88\% in alfalfa and rice, respectively).
Alfalfa datasets included heterozygous SNP loci, accounting for an average of 30\% of individuals per locus (Table~\ref{tab:descriptive_statistics}). Heterozygotes had an intermediate imputation accuracy ranging from 0\% with MNI to 70\% with KNNI. Results for each individual alfalfa population and the 12 individual rice chromosomes are similar to those discussed here (see Supplementary Figures S2 to S16). \\
Missing data did not affect the imputation accuracy substantially, with most algorithms showing a flat or almost flat response to increased missing rate. KNNI showed a decreasing imputation accuracy with increasing missing data, both in alfalfa and rice.

Beagle outperformed all other imputation methods in rice, making efficient use of marker position (Figure~\ref{fig:accuracy_rice}). When markers were randomly reshuffled (thus offsetting position information), though, general imputation methods (other than MNI) showed a higher imputation accuracy than Beagle. The accuracy loss with shuffled markers is exacerbated in alfalfa datasets, and in particular on heterozygous and minor homozygous classes.\\
Similarly, FILLIN resulted in a 5-10\% accuracy drop when marker positions were shuffled. FILLIN performances were in general lower than other methods due to the fraction of unimputed genotypes, here accounted as wrong imputation. FILLIN did not impute an average of 14.6\% missing data in Rice and practically the entirety of Alfalfa data. Even trying to relax FILLIN constraint parameters, it was impossible to obtain sensible results. Thus, FILLIN accuracy does not appear in Figure~\ref{fig:accuracy_alfalfa}.

When examining general imputation methods, resulting accuracies highlighted RFI and KNNI as the best ones. The two methods had comparable performances, without a preferable option over all experiments. SVDI followed, and MNI ranked last.\\
In Rice, for 70\% allowed missing rate datasets, the imputation accuracy of RFI was still greater than 90\%, while it dropped towards 75\% for KNNI and SVDI. As expected, MNI had an imputation accuracy of 100\% in the major homozygous class, but dropped to 0\% in the minor homozygous class.\\
In alfalfa KNNI showed the highest imputation accuracy over all classes (averaging 82.20\%) and in the heterozygous (66.02\%) and minor homozygous (14.68\%) genotype classes (Figure~\ref{fig:accuracy_alfalfa}). RFI and SVDI were second ranking. They did not differ much from MNI in the overall accuracy (MNI: 78.38\%, RFI: 79.50\%, SVDI: 77.35\%) and in the major (MNI: 88.68\%, RFI: 89.05\%, SVDI 85.81\%) homozygous class, but gave higher imputation accuracy in the heterozygous (MNI: 58.87\%, RFI: 61.75\%, SVDI 61.67\%). On minor homozygous class only SVDI resulted in some sporadic correct imputation, while RFI and MNI did not produce any sensible result (MNI: 0\%, RFI: 0.01\%, SVDI 4.46\%) classes.

>>>>

\subsection{Genotypes}
\label{sec:genotypes}  
GBS detected 32\,706, 40\,734 and 166\,418 SNP loci in Alfalfa-PV, Alfalfa-Med and Rice, respectively (Table~\ref{tab:descriptive_statistics}). The overall missing rate was $66.6\%$, $59.6\%$ and $53.4\%$ in the three datasets. Supplementary Figure S1 shows the density distribution of missing rate per marker in the complete datasets. The amount of missing marker genotypes varied with the threshold of maximum per-marker missing-rate allowed in the data (10\%, 20\%, 40\% and 70\%) and the proportion of artificial missing genotypes that were introduced (1\%, 5\%, 10\% and 20\%). Missing data point counts ranged from a minimum of 3\,878 in Alfalfa-PV with 10\% allowed and 1\% artificial missing genotypes, to a maximum of 5\,580\,616 in Rice with 70\% allowed and 20\% artificial missing genotypes. Over all datasets and allowed/artificial missing genotypes thresholds, the amount of missing data points to be imputed averaged 249\,405.\\
Minor Allele Frequency (MAF) was 0.172, 0.170 and 0.140 in Alfalfa-PV, Alfalfa-Med and Rice, respectively. The three datasets were therefore unbalanced with respect to genotype classes: Rice had the lowest MAF, while Alfalfa-PV and Alfalfa-Med had the smallest minor classes (minor homozygotes 3.5\% and 3.2\% of the total genotypes). Figure~\ref{fig:genotype_classes} reports the proportion of genotypes in each class in the complete datasets (100\%) and as a function of the maximum allowed missing rate per marker. The relative proportion of genotype classes remained approximately stable in rice. In contrast, the heterozygous class in alfalfa tended to become smaller with increasing proportion of allowed genotypes. This reflected the different GBS calling criteria for homozygous and heterozygous SNP: heterozygous SNPs required fewer overlapping reads to be called, while the implemented quality filters on alfalfa required a larger number of reads to call a homozygous locus. Therefore homozygous loci tended to have higher missing rates, and to
be included progressively more frequently with more relaxed thresholds on allowed missing level.
The average missing-rate and MAF in the 12 rice chromosomes were comparable to the whole-rice dataset.\\
Alignment on reference genomes assigned a position to 88.66\%, 57.54\% and 57.86\% of Rice, Alfalfa-Med and Alfalfa-PV markers, respectively, reflecting the difference between having a reference genome available for the exact same species (rice) or only for a close one (alfalfa).

dsahe different proportion of unaligned markers directly reflect the .

@@

In rice, the average imputation accuracy over all genotype classes was above 80\% for all methods. No significant difference was found between chromosomes. Thus average accuracy over all chromosomes is therefore presented (Figure~\ref{fig:accuracy_rice}). \\
The overall accuracy for alfalfa, averaged across the two datasets, was 12-25\% lower than rice, ranging between 66\% and 84\%. No significant differences were found between the two datasets. Thus average accuracy over the Alfalfa-Med and Alfalfa-PV is presented (Figure~\ref{fig:accuracy_alfalfa}).

The imputation accuracy varied across the different genotype classes, with the highest accuracy in the most common genotype (averaging 92.27\% and 96.47\% in alfalfa and rice, respectively), while the least common genotype class showed much lower accuracy (averaging 5.79\% and 69.88\% in alfalfa and rice, respectively).
Alfalfa datasets included heterozygous SNP loci, accounting for an average of 27\% of individuals per locus (Table~\ref{tab:descriptive_statistics}). Heterozygotes had an intermediate imputation accuracy ranging from 0\% with MNI to 70\% with KNNI. Results for each individual alfalfa population and the 12 individual rice chromosomes are similar to those discussed here (see Supplementary Figures S2 to S15). \\
Missing rate did not affect the imputation accuracy substantially, with most algorithms showing a flat or almost flat response to increased missing rate. KNNI showed a decreasing imputation accuracy with increasing missing rate, both in alfalfa and rice.

Beagle outperformed all other imputation methods in rice, making efficient use of marker position (Figure~\ref{fig:accuracy_rice}). When markers were randomly reshuffled (thus offsetting position information), though, general imputation methods (other than MNI) showed a higher imputation accuracy than Beagle. The accuracy loss with shuffled markers is exacerbated in alfalfa datasets, and in particular in heterozygous and minor homozygous genotypes.\\
Similarly, FILLIN resulted in a 5-10\% accuracy drop when marker positions were shuffled. FILLIN performances were in general lower than other methods due to the fraction of unimputed genotypes, here accounted as imputation errors. On average, FILLIN did not impute 14.6\% of missing data in Rice and practically the entirety of Alfalfa data. Even trying to relax FILLIN parameters, it was impossible to obtain sensible results in alfalfa. Thus, FILLIN accuracy results are not reported in Figure~\ref{fig:accuracy_alfalfa}.

When examining general imputation methods, resulting accuracies highlighted RFI and KNNI as the best ones. The two methods had comparable performances over all experiments. SVDI followed, and MNI ranked last.\\
In Rice, for 70\% allowed missing rate, the imputation accuracy of RFI was still higher than 90\%, while it dropped towards 85\% for KNNI and SVDI. As expected, MNI had an imputation accuracy of 100\% in the major homozygous class, but dropped to 0\% in the minor homozygous class.\\
In alfalfa KNNI showed the highest imputation accuracy over all classes (averaging 82.20\%) and in the heterozygous (66.02\%) and minor homozygous (14.68\%) genotype classes (Figure~\ref{fig:accuracy_alfalfa}). RFI and SVDI were second ranking. They did not differ much from MNI in the overall accuracy (MNI: 78.38\%, RFI: 79.50\%, SVDI: 77.35\%) and in the major (MNI: 88.68\%, RFI: 89.05\%, SVDI 85.81\%) homozygous class, but gave higher imputation accuracy in the heterozygous (MNI: 58.87\%, RFI: 61.75\%, SVDI 61.67\%). In the minor homozygous class only SVDI resulted in some sporadic correct imputation, while RFI and MNI did not produce any sensible result (MNI: 0\%, RFI: 0.01\%, SVDI 4.46\%) classes.
